Sparse Function Space Representation of Neural Networks for Exploration and Retention

### 2023-05-10 paper/main.tex
starting to fill out background

\textbf{Dual GP} A Gaussian process is a prior over functions $f \sim \GP(0, \kappa)$. The prior is characterized by a covariance kernel function $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors. Once combined with a data set ${\mathcal{D}_n = \{(\vx_i, y_i)\}_{i=1}^n}$ with $n$ examples, each modeled by the likelihood $p(y_i \mid f_i)$ at function $f_i = f(\vx_i)$, we can form a posterior using multivariate conditioning.

Sparse Gaussian processes \citep{hensman2013gaussian} are a principled way to scale Gaussian processes to large datasets by changing the complexity from $O(n^3)$ where n is number of data points to $O(m^3)$ with $m$ the number of inducing points, where typically $m \ll n$. The key to the complexity saving is to use a sparse distribution $q(\vu; \vm, \MV)$, where $\vm \in \R^{m}$ and $\MV \in \R^{m \times m}$. The distribution at all over points is then defined as $q_{\vu}(f) = \int p(f \mid \vu) q(\vu) d\vu$.

dual parameterization states that we can decompose the optimal sparse Gaussian process posterior as follows:
\begin{align}
   \myexpect_{q_{\vu}(f)}[f_i] &= \vkzs^{\T} \valpha_{\vu} , \\
   \textrm{Var}_{q_{\vu}(f)}[f_i]  &= \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \vbeta_{\vu})^{-1} ]\vkzs.
\end{align}
The dual parameters consist of an $m$-length vector $\valpha_\vu$ and $m\times m$-size matrix $\vbeta_\vu$, defined as
\begin{equation} 
\begin{split}
   \vlambda & = [ \MKzz + \vbeta_{\vu}]\valpha_{\vu}  = \sum_{i \in \mathcal{D}_{\textrm{old}} \cup \mathcal{D}_{\textrm{new}}}  \vkzi \, \tilde{\lambda}_{i}, \\
   \vbeta_{\vu} &= \sum_{i \in \mathcal{D}_{\textrm{old}} \cup \mathcal{D}_{\textrm{new}}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,
\label{eq:dual_sparse}
\end{split}
\end{equation}
where $\tilde{\lambda}_i = \alpha_i - \myexpect_{q_{\vu}(f)}[f_i] \beta_i$ with $(\hat{\alpha}_i, \hat{\beta}_i)$ defined as follows,
\begin{equation}
\begin{aligned}
   \alpha_i &= \myexpect_{q_{\vu}(f_i)} [\nabla_{f_i}\log p(y_i \mid f_i)],\\
   \beta_i &= \myexpect_{q_{\vu}(f_i)} [ -\nabla^2_{f_i}\log p(y_i \mid f_i)],
\end{aligned} \label{eq:svgp_dual}
\end{equation}

>>>>

\textbf{Dual Gaussian Process} A Gaussian process (GP) is a prior over functions $f \sim \GP(0, \kappa)$. The prior is characterized by a covariance kernel function $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors typically in $\R^{\mathrm{D}}$. Once combined with a data set ${\mathcal{D}_n = \{(\vx_i, y_i)\}_{i=1}^n}$ with $n$ examples, each modeled by the likelihood $p(y_i \mid f_i)$ at function $f_i = f(\vx_i)$, we can form a posterior proccess using multivariate conditioning. The drawbacks to GPs are their lack of scaling to large datasets and non-tractability with non-Gaussian likelihoods.

Sparse Gaussian processes \citep{hensman2013gaussian} are a principled way to scale Gaussian processes to large datasets by changing the complexity from $O(n^3)$ to $O(m^3)$ with $m$ the number of inducing points, and typically $m \ll n$. The key insight is to use a sparse approximating distribution $q(\vu; \vm, \MV)$, where $\vm \in \R^{m}$ and $\MV \in \R^{m \times m}$, and find its parameters through maximizing the evidence lower bound (ELBO). The distribution at all over points is then defined as $q_{\vu}(f) = \int p(f \mid \vu) q(\vu) d\vu$. The dual parameterization \citet{adam2021dual} states that we can decompose the optimal sparse Gaussian process posterior as follows:
\begin{equation}
   \myexpect_{q_{\vu}(f)}[f_i] = \vkzs^{\T} \valpha , \quad 
   \textrm{Var}_{q_{\vu}(f)}[f_i]  = \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \vbeta)^{-1} ]\vkzs.
\end{equation}
The dual parameters consist of an $m$-length vector $\valpha$ and $m\times m$-size matrix $\vbeta$, defined as
\begin{equation} 
   \vlambda = [ \MKzz + \vbeta]\valpha  = \sum_{i \in \mathcal{D}_{n}}  \vkzi \, \tilde{\lambda}_{i}, \quad
   \vbeta = \sum_{i \in \mathcal{D}_{n}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,
\label{eq:dual_sparse}
\end{equation}

where $\tilde{\lambda}_i = \alpha_i - \myexpect_{q_{\vu}(f)}[f_i] \beta_i$ with $(\hat{\alpha}_i, \hat{\beta}_i)$ defined as follows,
\begin{equation}
 \alpha_i = \myexpect_{q_{\vu}(f_i)} [\nabla_{f_i}\log p(y_i \mid f_i)], \quad
   \beta_i = \myexpect_{q_{\vu}(f_i)} [ -\nabla^{2}_{f_i}\log p(y_i \mid f_i)],
\label{eq:svgp_dual}
\end{equation}
